### 2018-10-18 EMIN-2018/article/emin2108.tex
save

\documentclass[a4paper]{panl}

\usepackage{cite}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{epsfig}
\usepackage{multirow}

\originalTeX
%\russianTeX
\begin{document}
% Journal sections (see http://pkp.jinr.ru/index.php/PEPAN_LETTERS/about/editorialPolicies#focusAndScope)
\issuearea{Physics of Elementary Particles and Atomic Nuclei. Theory}
% or in Russian
%\issuearea{ФИЗИКА ЭЛЕМЕНТАРНЫХ ЧАСТИЦ И АТОМНОГО ЯДРА. ТЕОРИЯ}

\title{Chemical composition analysis for X-ray transport container scans.  \\ Анализ химического состава при сканировании транспортных контейнеров гамма-излучением}
\maketitle
\authors{A. Zelenaya$^{~a}$, M. Zelenyi$^{~a,b,}$\footnote{E-mail: mihail.zelenyy@phystech.edu}, A.A.Turinge$^{~a}$,  V.G. Nedorezov$^{~a,}$\footnote{E-mail: vladimir@cpc.inr.ac.ru}}
\from{$^{a}$\,Institute of Nuclear Research of RAS}
\vspace{-3mm}
\from{$^{b}$\,Moscow Institute of Physical and Technology  (SU)}

\begin{abstract}
% Russian translation of the abstract
It is important for national security to control the movement of dangerous or strategically cargo such as explosives, radioactive materials, rare and precious metals. This control can be provided by scanning transport containers by gamma rays.
In this report the existing technique for scanning (dual energy method) is considered and the alternative method based on measuring the energy distribution of gamma rays is proposed. For estimation perspectives of the proposing method, the  corresponding simulation was conducted by using the GEANT4 toolkit. The example of the algorithm of reconstruction the chemical composition of the scanned object is also considered. In addition the experiment for estimation energy resolution of the detector based on a scintillation crystal BGO and SiPM was carried out.\\
\vspace{0.2cm}

Для обеспечения национальной безопасности важен контроль перемещения опасных или стратегически важных грузов, таких как взрывчатые вещества, радиоактивные материалы, редкие и драгоценные металлы. Проводить такой контроль можно сканирую содержимое   транспортных контейнеров гамма излучением. В данной работе рассмотрена существующая методика дуальных энергий и предложен альтернативный способ, основанный на измерении энергетического распределения гамма-квантов. Для оценки было проведено моделирование с помощью транспортного кода GEANT4.  Также выполнен эксперимент по измерению энергетического разрешения детектора на основе сцинтиллирующего кристалла BGO и кремневого фотоумножителя. 
\end{abstract}
\vspace*{6pt}

\noindent
PACS: 02.70.$-$c; 23.20.Nx; 32.90.$+$a

\newpage
\label{sec:intro}
\section*{Introduction}
Geant4~\cite{ALLISON2016186, spirin}
\section*{Dual energy method}

\begin{figure}[t]
\begin{center}
    % Сделать подписи к отдельным картинкам
\includegraphics[width=60mm]{figures/Attenuation.pdf} 
\includegraphics[width=60mm]{figures/Bremsstrahlung.pdf}  
\vspace{-3mm}
\caption{a)Attenuation curve b) Bremsstrahlung spectrum by $10 MeV$ electron}
\end{center}
\labelf{fig01}
\vspace{-5mm}
\end{figure}

            $$
T(E_0, t, Z) = \frac{\int \limits_0^{E_0} S(E_0, E) \exp(-\mu(E,Z)\times t)~dE)}{\int \limits_0^{E_0} S(E_0, E)~dE},
$$
where
$T$ -  transmittance,
$S(E_0, E)$ - response function,
$\mu(E,Z)$ - attenuation,
$t$ -  optical thickness,
$E_0$ -  up-limit energy of bremsstrahlung,
$E$ - energy of gamma ray,
$Z$ - charge of nuclei.
We can't defined Z if optical thickness $t$ is unknown.
But we can use two electron beams with different energy which give gamma rays with up-limit energy $E^{(1)}_0$ and $E^{(2)}_0$.
 Then we can get Z as a result of minimizing this function:
    $$
    F(z) = \frac{|t(E^{(1)}_0,z) - t(E^{(2)}_0,z)|}{t(E^{(1)}_0,z)} \to min
    $$
This technique allows to determine scan object as a one from four possible groups: $Z_{eff} \sim 5$, $Z_{eff} \sim 13$, $Z_{eff} \sim 26$, $Z_{eff} \sim 82$.\\
Disadvantages of dual energy method
    \begin{itemize}
        \item It is too difficult to irradiate the target with beams with different energy.
        \item Low efficiency for target which contains elements with strongly different charges.
    \end{itemize}
Our method
    \begin{itemize}
        \item Use only one electron beam with energy $E = 10 MeV$.
        \item Measure not only the space distribution, but also the energy of gamma rays.
    \end{itemize}

\section*{Simulation}
\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=120mm]{figures/yed_schema_1.pdf}
        \vspace{-3mm}
        \caption{a) }
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}
\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        % Переделать картинку от Туринге
        \includegraphics[width=60mm]{figures/Sword.pdf} 
        \includegraphics[width=60mm]{figures/Sword.pdf}  
        \vspace{-3mm}
        \caption{a) b)}
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}
\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        % Переделать картинку от Туринге
        \includegraphics[width=60mm]{figures/Difference.pdf} 
        \includegraphics[width=60mm]{figures/Difference.pdf}  
        \vspace{-3mm}
        \caption{a) b)}
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}

\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        % Переделать картинку от Туринге
        \includegraphics[width=60mm]{figures/UranCube1.pdf} 
        \includegraphics[width=60mm]{figures/UranCube2.pdf}  
        \vspace{-3mm}
        \caption{a) b)}
    \end{center}
    \labelf{fig01}
    \vspace{-5mm}
\end{figure}

\section*{Thickness reconstruction}
\begin{wrapfigure}[8]{r}{0.6\linewidth} 
    % Сделать подписи к отдельным картинкам
    \includegraphics[width=\linewidth]{figures/yed_schema_2.pdf}
    \vspace{-3mm}
    \caption{ Thickness reconstruction}
    \labelf{schema2}
    \vspace{-5mm}
\end{wrapfigure}
Мы будем рассматривать одномерный случай, когда гамма-кванты проходят скозь стопку из нескольких материалов с фиксированной полной толщиной и нам нужно восстановить толщины отдельных материалов (см. схему ~\ref{scema2}). Мы будем использовать простую физическую модель в которой ослабление задается  следующей формулой: 
\begin{equation}
\label{eq:gamma}
\frac{N(E)}{N_0(E)} = \exp(-\sum_i \Sigma^{mean}_i(E)x_i),
\end{equation}
where $x_i$ --- thickness of the $i$-layer, $\Sigma^{mean}_i$ --- mean cross-section for group of materials with близкими зарядами ядра, $N,~N_0$ --- the number of gamma.При этом мы не будем учитывать множественное рассеяние и наличие аннигиляционной линии. Мы считаем что полная толщина известна и для восстановления толщин отдельных слоев мы будем использовать метод наименьших квадратов, то есть минимизировать такую сумму:
\begin{equation}
\sum_E(\ln \frac{N(E)}{N_0(E)} + \sum_i \Sigma^{mean}_i(E)x_i))^2 \to min
\end{equation}
Далее приведен пример работы алгоритма. We consider the object which has 3 layers of aluminium, iron and lead.The figure ~\ref{rec:ex} shows a contribution of every reconstructed material in summary attenuation. The table ~\ref{tab:rec} contains the results of reconstruction.
\begin{figure}[t]
    \begin{center}
        % Сделать подписи к отдельным картинкам
        \includegraphics[width=60mm]{figures/reconstruction.pdf} 
        \includegraphics[width=60mm]{figures/relError.pdf}  
        \vspace{-3mm}
        \caption{a) b)Distribution of reconstruction errors for various numerical experiments}
    \end{center}
    \labelf{rec:ex}
    \vspace{-5mm}
\end{figure}

\begin{table}
\begin{center}
        \begin{tabular}[c]{|c|c|c|}
        \hline 
        Material & Real, cm & Reconstructed, cm \\ 
        \hline 
        Al & 20 & 19.6 \\ 
        \hline 
        Fe & 40 & 41.6 \\ 
        \hline 
        Pb & 30 & 28.7 \\ 
        \hline 
    \end{tabular}
\end{center}
\caption{Example of thickness reconstruction}
    \label{tab:rec}
\end{table}

\section*{Experiment}
\begin{wrapfigure}[20]{r}{0.5\linewidth} 
    % Сделать подписи к отдельным картинкам
    \includegraphics[width=\linewidth]{figures/yed.pdf}  
    \vspace{-3mm}
    \caption{Schema of experiment}
    \labelf{pic:experiment}
    \vspace{-5mm}
\end{wrapfigure}
Кроме моделирования, доктором Губером и доктором Ивашкины, проведены измерения энергического разрешения детектора  для регистрации гамма излучения. Схема эксперимента приведена на рисунке~\ref{pic:experiment}. В работе использовались сцинтиллятор BGO размером 10х30х100мм и фотодетектор ArrayC-60035-4P. Проводились измерения спектров от источников $\beta$– излучения (Na-22 и Cs-137). Энергии излучения 0.511 МэВ и 1.275 МэВ для Na и 0.662 МэВ для Cs. В качестве фотодетектора использовалась матрица из четырех фотодиодов ArrayC-60035-4P с размером фотодиода 6х6мм, снабженная индивидуальным предусилителем с коэффициентом усиления 150. В процессе работы измерялся суммарный сигнал с двух фотодиодов матрицы. Сигнал c матрицы подавался на усилитель (ORTEC 579) и затем поступал на входной канал АЦП (CAEN DT5742) и на дискриминатор (CAEN mod.224), логический сигнал которого служил триггером в системе. В отсутствие источника шкала АЦП была прокалибрована в абсолютных единицах – числе фотоэлектронов. Результат измерения энергитического разрешения приведен в таблице ~\ref{tab:ex}

The experiment was conducted by Dr. Guber and Dr. Ivashkin.
The sum of signals from 2 photodiode\\
Noise threshold: 100 KeV\\
\begin{table}
    \begin{center} 
        \begin{tabular}[c]{|c|c|c|}

            \hline 
            Source & Energy, MeV & Sigma/Mean \\
            \hline 
            $^{22}Na$&0.511 & 14.7\%  \\ 
            \hline 
            $^{137}Cs$&0.662 & 19\%\\ 
            \hline 
            $^{22}Na$& 1.275 & 13\% \\
            \hline 
        \end{tabular} 
    \end{center}
\caption{Energy resolution of detector}
\label{tab:ex}
\end{table}

\section*{Conclusion}
% Выводы сделать более твердыми и акцентирванными
Results:   
    \begin{enumerate}
        \item The measurement of the gamma ray spectrum allows to identify cargo belonging to the group of materials with certain $Z_{eff}$.
        \item Also it allows to define the thickness of layers from different elements with the accuracy about 25\%.
        \item The energy resolution of the detector based on a BGO scintillator was studied.  For the photodetector with full array of pixels the energy resolution is expected about 10\%.
    \end{enumerate}
Plans and perspectives
    With financial support can be developed:
    \begin{enumerate}
        \item The program which checks cargo of a transport container  for compliance cargo manifest
        \item The algorithm for the 3D gamma-tomography.
    \end{enumerate}

This work is supported by the Ministry of Education and Science of the Russian Federation under the contract No. 3.3008.2017/PP.

%============================= Fig. 1 ================================
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=127mm]{fig_1.eps} 
%%\includegraphics[width=127mm]{fig_1.eps} 
%\vspace{-3mm}
%\caption{Figure caption}
%\end{center}
%\labelf{fig01}
%\vspace{-5mm}
%\end{figure}
%============================= Fig. 1 ================================

%\nocite{*}
\bibliographystyle{pepan}
\bibliography{references.bib}

\end{document}