### 2018-12-05 MIPT-2018/satellite/main.tex

\documentclass[11pt,twoside,openany,a4paper,final]{book}

\usepackage{cmap}
\usepackage{mathtext}
\usepackage[T2A]{fontenc}
\usepackage[cp1251]{inputenc}
\usepackage[russian,english]{babel}

\usepackage{indentfirst}
\usepackage{misccorr}
\usepackage[font=small, labelsep=period]{caption}
\usepackage{cite}
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{nicefrac}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{multirow}

\usepackage[pdftex, left=1in, right=1in, top=1in, bottom=2cm]{geometry}

\usepackage{mycommon}
\usepackage{style}

\newcommand{\issuemark}{ТРУДЫ МФТИ.~--- 2015.~--- Том 7, \textnumero~3}
\setcounter{MaxMatrixCols}{35}

\sloppy

\begin{document}

\selectlanguage{russian}

\author[1]{И1.\,О1.}{Фамилия1}
\author[2]{И2.\,О2.}{Фамилия2}

\affil[1]{Место работы 1}
\affil[2]{Место работы 2}

\title{Название статьи}

\inputarticle[xxx.yyy]{familia}{И1.\,О1. Фамилия1, И2.\,О2. Фамилия2}

\selectlanguage{russian}

\include{authors}

\include{links}

\end{document}